---
title: "D213-Task01-PartF"
author: "Raquel Ocasio"
date: "2024-03-14"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Part A1
Can daily revenue for the organization be predicted for the next year?

## Part A2
The goal of the data analysis is to use time series modeling to predict the daily revenue of the organization for the next year.

## Part B
Time series models rely on several assumptions to ensure their validity and reliability. Fundamental among these is stationarity, which asserts constancy in statistical properties over time, either strictly or with respect to mean. Autocorrelation, indicating correlation between a time series and its past values, is also crucial. Homoscedasticity assumes a consistent variance of residuals, while normality of residuals and independence between them facilitate effective statistical inference. Additionally, the absence of seasonality, or proper adjustment for it, is vital, as is addressing outliers that may disrupt model performance. Adequate sample size is required to accurately capture the complexity of the time series. 
```{r,results=FALSE, warning=FALSE, message=FALSE}
# Install necessary packages
library(forecast)
library(tseries)
library(ggplot2)
library(seasonal)
library(caTools)
library(astsa)
library(graphics)

# Read data from CSV file into a dataframe
telecodata <- read.csv("teleco_time_series.csv")

# Data preparation
# show dimensions, structure, and sample data
str(telecodata)
# check for NA values
colSums(is.na(telecodata))
# check for null values
is.null(telecodata)
# check for duplicates in Day column
anyDuplicated((telecodata$Day))
# save prepared dataset to CSV
write.csv(telecodata, "telecodata_fulldataset.csv", row.names=FALSE)

# C1
# Create a time series plot
ggplot(telecodata, aes(x = Day, y = Revenue)) +
  geom_line() +
  labs(title = "Time Series Plot of Revenue",
       x = "Day",
       y = "Revenue in Millions")
```

## Part C2
The time step formatting of the realization is a day. The length of the sequence is 731 days, and there are no gaps in measurement.
```{r,results=FALSE, warning=FALSE, message=FALSE}
# C2
# verify dataframe
is.data.frame(telecodata)
head(telecodata)
# convert Revenue column to vector
telecodata_vector <- telecodata[['Revenue']]
# verify vector
is.vector(telecodata_vector)
# convert to time series
telecodata_ts <- ts(telecodata_vector, frequency = 731/24)
# verify time series object
is.ts(telecodata_ts)
```

## Part C3
The Augmented Dickey-Fuller test was applied to the data to determine stationarity of the time series.  The test yielded a p-value of 0.02431, which is less than the p-value of 0.05. Therefore, the time series is likely stationary.
```{r,results=FALSE, warning=FALSE, message=FALSE}
# C3
# ADFuller test for stationarity
adf_result <- adf.test(telecodata_ts)

# Print ADFuller test results
print(adf_result)

# Additional information based on ADFuller test
if (adf_result$p.value < 0.05) {
  cat("The time series is likely stationary (p-value < 0.05)\n")
} else {
  cat("The time series is likely non-stationary (p-value >= 0.05)\n")
}
```

## Part C4
The data was prepared for analysis by checking for: NA values, null values, duplicates, stationarity, and splitting the data into test and training sets.


## Part C5
See attached files.
```{r,results=FALSE, warning=FALSE, message=FALSE}
# C5
#make this example reproducible
set.seed(1)

#use 70% of dataset as training set and 30% as test set
sample <- sample.split(telecodata$Day, SplitRatio = 0.7)
train  <- subset(telecodata_vector, sample == TRUE)
write.csv(train, "train_data.csv", row.names=FALSE)
test   <- subset(telecodata_vector, sample == FALSE)
write.csv(train, "test_data.csv", row.names=FALSE)

# convert train set to time series
telecodatatrain_ts <- ts(train, frequency = 731/24)
# verify time series object
is.ts(telecodatatrain_ts)
# convert test set to time series
telecodatatest_ts <- ts(test, frequency = 731/24)
# verify time series object
is.ts(telecodatatest_ts)
```


## Part D1
Plot for seasonal component
```{r,results=FALSE, warning=FALSE, message=FALSE}
# D1
# Decomposed time series
decomposition <- decompose(telecodata_ts)

# plot seasonality
plot(decomposition$seasonal, main = "Seasonal Plot")

# plot trend
plot(decomposition$trend, main = "Trend Plot")

# Calculate and plot autocorrelation
acf_result <- acf(telecodata$Revenue, lag.max = 30) 
plot(acf_result, main = "Autocorrelation Function for Original Time Series")

# plot spectral density
specden <- spectrum(telecodata_ts)
# Plot the spectrum with a title
plot(specden, main = "Spectral Density Plot")

# plot decomposed time series
stl <- stl(telecodata_ts, s.window="period")
autoplot(stl, main = "Decomposed Time Series Plot")

# plot residuals
# Calculate the residuals
residuals <- telecodata$Revenue - (decomposition$trend + decomposition$seasonal + decomposition$random)

# Plot
ggplot(telecodata, aes(x = Day, y = residuals)) +
  geom_line(color = "orange") +
  labs(title = "Residuals of the Time Series",
       x = "Day",
       y = "Residuals") +
  theme_minimal()
```

## Part D2
```{r,results=FALSE, warning=FALSE, message=FALSE}
# D2
# automate model selection
auto.arima(telecodata_ts, seasonal = T)

# fit model
fit <- Arima(telecodata_ts, order = c(1,1,0))

# run summary statistics on model
summary(fit)

# review residual pattern
checkresiduals(fit)
```

## Part D3
```{r,results=FALSE, warning=FALSE, message=FALSE}
# D3
# forecast using model
forecast <- forecast(fit, h = 183)

# visualize forecast based on training data
plot(forecast)

# add testing data points for comparison
lines(test)

# visualize p-values for Ljung-Box statistics
Box.test(forecast, lag = 1, type = c("Ljung-Box"))
```

## Part D4
See attached code file.


## Part D5
See attached code file.


## Part E1
The ARIMA model was selected based on the output of the auto.ARIMA() function. 
The forecast's prediction interval is set at 1 day, and our dataset consists of 2 years of daily revenue information. Consequently, the ARIMA model is utilized to uncover correlations and seasonality within the data, enabling the prediction of revenue at a daily interval. 
Given a dataset spanning 2 years of daily revenue, TM has the capability to forecast future revenue for up to 1 year. Predictions made within the first year are expected to be more accurate. For more reliable long-term predictions, a larger historical dataset is necessary. 
The model was evaluated automatically by the auto.ARIMA() function, which produced the model with the best AIC. The standard error (s.e.) was used as the error metric. The model produced by the  auto.ARIMA() function produced a low s.e. value of 0.0327.


## Part E2
```{r,results=FALSE, warning=FALSE, message=FALSE}
# E2
# forecast with new model using test data converted to time series
sarima.for(telecodatatest_ts, n.ahead = 731, 1,1,0, plot.all = T, xlab="Months", main = "Revenue Forecast vs. Test Data")
# need to start line for test data at specific X coordinate
#valueymod <- c(rep(16, 350), telecodatatest_ts)
#lines(telecodatatest_ts, valueymod, type="l")

# annotations
text(5,32,labels="Test Data")
arrows(5,29,5,15)
text(15,70,labels="Forecast")
arrows(15,66,15,34)
text(27,10, labels="Confidence Intervals")
arrows(25,12,25,35)
arrows(26,12,26,50)
arrows(27,12,27,69)
arrows(28,12,28,85)
```

